\input{header}

\section{Introduction}
\begin{frame}
\frametitle{Covariance: A Motivating Example}
%

\begin{center}
\colorbox{green!10}{
$\textnormal{Gaussian pdf} = 
    \color{m1}(2\pi{})^{\!-\frac{n}{2}}\det(\mSigma)^{\!-\frac{1}{2}}
           e^{(\mx-\mu)^\top{}\mSigma^{-1}(\mx-\mu)}$
}
\end{center}
%
Tikhonov regularized maximum log-likelihood for learning the covariance
$\color{m1}\mSigma$ given an empirical covariance $\color{m1}\mS$ is:
%
\begin{center}
%
\colorbox{green!10}{
%
$\color{m1}
f(\mSigma)=-\logdet(\mSigma)-\trace(\mS\mSigma^{-1})-\|\mSigma^{-1}\|_F^2.$
%
}
%
\end{center}
%
\begin{itemize}
%
\item To optimize $\color{m1}f(\mSigma)$, using first-order methods we need 
      $\color{m1}f'(\mSigma)$.
%
\item To use second-order methods, we also need $\color{m1}f''(\mSigma)$.
%
\end{itemize}
%
\begin{center}
%
What are $\color{m1}f'(\mSigma)$ and $\color{m1}f''(\mSigma)$?
%
\end{center}
%
\end{frame}

%
\begin{frame}
\frametitle{Objective}
\framesubtitle{Theoretical Aspect}
%
\begin{center}
%
Differentiation of matrix functions is still not well understood.
%
\end{center}
%
\begin{itemize}
%
\item No simple ``calculus'' for computing matrix derivatives.
%
\item Matrix derivatives of simple functions can be complicated.
%
\end{itemize}
%
\begin{center}
%
\textcolor{blue}{Build on known results to define a matrix-derivative
calculus.}
%
\end{center}
%
\end{frame}

\begin{frame}
\frametitle{Objective}
\framesubtitle{Software Aspect}
%
\begin{center}
\textcolor{blue}{Matrix differentiation made simple for humans and computers}
\end{center}
%
\begin{itemize}
\item Symbolic differentiation should be made simple.
\item Optimizing matrix valued functions should be efficient:
  \begin{itemize}
  \item Both computation and storage.
  \end{itemize}
\end{itemize}

%
\begin{center}
We are developing Automatic Matrix Differentiation (AMD), an open-source \Cpp{}
library to accomplish this.  The code can be downloaded from
\textcolor{blue}{\url{https://github.com/pkambadu/AMD}}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Description}
\begin{center}
\textcolor{blue}{Re-design AMD from ground up}
\end{center}

\begin{itemize}
\item Previous implementation led to key insights.
\item Rewriting from scratch better than fixing existing design.
    \begin{itemize}
    \item Easy creation of matrix-matrix functions.
    \item Optimization of abstract syntax trees.
    \item Advanced logging and exception handling.
    \item Parallel evaluation on several matrix types.
    \item Python bindings.
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Expected \Cpp{} Interface}

\begin{lstlisting}[style=basic]
@{\color{red} /** Associate matrices with symbols */}@
std::map<std::string, boost::shared_ptr<MatrixType> > symbolTable;
symbolTable["A"] = ...; symbolTable["X"] = ...;

@{\color{red} /** Create an expression and evaluate it */}@
AMD::Expression fX("trace(A*(I+X)");
AMD::evaluate(fX, symbolTable);

@{\color{red} /** Evaluate the derivative */}@
AMD::Expression fX_prime = AMD::gradient(fX);
AMD::evaluate(fX_prime, symbolTable);
\end{lstlisting}
\end{frame}

\begin{frame}
\frametitle{Creating Expression Trees}
%
\begin{center}
\textcolor{blue}{Accept a well-formed matrix op-string and return an AST"}
\end{center}
%
\begin{itemize}
\item Use \texttt{Boost.Spirit}.
\item Matrix-matrix functions: +, -, *, .*, ',\textunderscore{}
\item Scalar-Matrix functions: \code{trace}, \code{logdet}.
\item Obey precedence and associativity (include parantheses).
\end{itemize}
%
\end{frame}

\begin{frame}
\frametitle{Optimizing Expression Trees}
%
\begin{center}
\textcolor{blue}{Reduce computational complexity by rewriting the AST}
\end{center}
%
\begin{itemize}
\item Use all known matrix identities.
\item \code{(A')' = A}, \code{(A_)_ = A}
\item \code{I*A = A*I = A}, \code{0+A = A+0 = A}
\item Common sub-expression elimination
\item $\ldots{}$
\end{itemize}
%
\end{frame}

\begin{frame}
\frametitle{Derivatives Of Scalar-Matrix Functions}
%
\begin{center}
\textcolor{blue}{Compute the derivative using reverse mode}
\end{center}
%
\begin{itemize}
\item Apply the rules from the AMD paper.
\item Apply optimizations post-derivation.
\end{itemize}
%
\end{frame}

\begin{frame}
\frametitle{Evaluation Expression Trees}
%
\begin{center}
\textcolor{blue}{Evaluate the AST in parallel}
\end{center}
%
\begin{itemize}
\item Identify parallelism (topological sort).
\item Evaluation on a symbol table.
\end{itemize}
%
\end{frame}

\begin{frame}[fragile]
\frametitle{Operator Associativity}
%
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Operator & Associativity & Priority \\ \hline
$( )$    &      ---      &    1  \\ \hline
$',\_$    &      left     &    2  \\ \hline
unary $+,-$ &   left     &    3  \\ \hline
$*,.*$    &     left     &    4  \\ \hline
$+,-$     &     left     &    5  \\ \hline
\end{tabular}
\end{center}
%
\begin{itemize}
\item \hyperref[http://www.boost.org/doc/libs/1_57_0/libs/spirit/doc/html/index.html]{Boost.Spirit}
\item References for calculator:
    \begin{itemize}
    \item \hyperref[http://stackoverflow.com/questions/13056893/building-a-custom-expression-tree-in-spiritqi-without-utree-or-boostvariant/13060402]{Building a Custom Expression Tree in Spirit:Qi}
    \item \hyperref[http://stackoverflow.com/questions/8464969/boostspirit-expression-parser/8468822]{Boost::Spirit Expression Parser}
    \item \hyperref[http://stackoverflow.com/questions/9453712/compilation-error-with-a-boostspirit-parser/9458029]{Compilation error with a boost::spirit parser}
    \end{itemize}
\end{itemize}
%
\end{frame}

\input{footer}
