\section{Software Contributions}

\begin{frame}
\frametitle{AMD Software}
\framesubtitle{Overview}
%
\begin{itemize}
\item Implemented using templates and operator over-loading.
\item Provides both \alert{symbolic} and \alert{algorithmic} differentiation.
\begin{itemize}
\item Symbolic differentiation gives results as Matlab code.
\item Algorithmic differentiation:
  \begin{itemize}
  \item Elemental (dense, multi-threaded, distributed).
  \item Eigen (dense and sparse, multi-threaded).
  \item Easy to extend to other matrix packages.
  \end{itemize}
\end{itemize}
\item Scalar-matrix: {\color{m1}trace} and {\color{m1}log-determinant}.
\item Matrix-matrix:
  $\color{m1}+$, $\color{m1}-$, $\color{m1}*$, $\color{m1}.*$, 
                              {\color{m1}transpose} and {\color{m1}inverse}.
\item Two ways to compute symbolic differentiation: 
  \begin{itemize}
  \item \texttt{C++} program or on the command line.
  \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{AMD Software}
\framesubtitle{Example AMD Program}
Define a Scalar-Matrix function $\color{m1} \trace(\mA\mX\mB\mX)$
\footnotesize
\begin{lstlisting}[style=basic]
  @{\color{red} /** Create a variable X and an identity function */}@
  symbolic_matrix_type X("X", ROW, COL);
  SymbolicMMFunc fX(X, false);
  @{\color{red} 
    /** Create constants A,B and corresponding constant functions */}@
  symbolic_matrix_type A("A", ROW, COL);
  symbolic_matrix_type B("B", ROW, COL);
  SymbolicMMFunc fA(A, true);
  SymbolicMMFunc fB(B, true);
  @{\color{red} /** Scalar-matrix function placeholder */}@
  SymbolicSMFunc func; 
  func = trace(fA*fX*fB*fX);
  std::cout << "Function Value:   " << func.functionVal.getString();
  std::cout << "Derivative Value: " << func.derivativeVal.getString();
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]
\frametitle{AMD Software}
\framesubtitle{Example AMD Program}
%
The resulting derivative is:
\begin{lstlisting}[style=basic]
$ ./SymbolicSample.exe
Function Value:   trace(((A*X)*B)*X)
Derivative Value: (((B*X)*A)'+((A*X)*B)')
\end{lstlisting}
%
\begin{center}OR\end{center}
%
Use the symbolic calculator:
\begin{lstlisting}[style=basic]
$ ./SymbolicCalculator.exe 'trace(((A*X)*B)*X)'
Function Value: trace(((A*X)*B)*X)
Derivative Value: (((B*X)*A)'+((A*X)*B)')
\end{lstlisting}
%
\end{frame}

\begin{frame}
\frametitle{AMD Software}
\framesubtitle{In Progress: Program Optimization}

\begin{itemize}
\item Common sub-expression elimination.
\item Repeated application of matrix and direct-matrix identities.
  \begin{itemize}
  \item $\color{m1}\mI\times{}\mX=\mX$, $\color{m1}\mX+0=\mX$,
        $\color{m1}(\mA^\top)^\top = \mA$, etc.
  \item $\color{m1}(\mI\otimes{}\mI)\myvec(\mI)=\mI$, 
        $\color{m1}(\mA\boxtimes{}\mB)^\top=\mB^\top\boxtimes\mA^\top$, 
        etc.
  \end{itemize}
\item Critical in providing performance guarantees.
\item Students working on this:
\begin{itemize}
\item Wuwei Zhang (Purdue University) 
\item Allan Sepucaia (University of Campinas).
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{AMD Software}
\framesubtitle{In Progress: Python Interface}
%
\begin{lstlisting}[style=basic]
$ python 
>>>import AMD
>>>X = AMD.symMat('X')
>>>fX = AMD.scalarMatrixFunc(X, 1);
@\dots{}@
\end{lstlisting}
%
\begin{itemize}
\item Easy to use and prototype.
\item Has a large user-base in machine learning.
\item Students working on this:
\begin{itemize}
\item Anna Romanova (UCSD) 
\item Gabriel Borges (University of Campinas).
\end{itemize}
\end{itemize}
\end{frame}
